Assessing Population Density and Crime Patterns in Toronto Neighborhoods (2014–2023)

### 2024-09-27 paper/paper.qmd
Revise typos.
Revise some typos.

---
title: "Assessing Population Density and Crime Patterns in Toronto Neighborhoods (2014–2023)"
author: 
  - Yongqi Liu
thanks: "Code and data are available at:https://github.com/Cassieliu77/Crime-Patterns-and-Population-Density.git"
date: 27 September 2024
date-format: long
toc: true 
abstract: "This paper examines the relationship between population density and crime distribution across Toronto neighborhoods from 2014 to 2023. The analysis reveals that high-density areas experience higher crime rates, particularly for auto theft and assault, with socioeconomic factors further influencing crime patterns. Based on these findings, we propose targeted interventions and offer data-driven suggestions for policymakers and urban planners to enhance crime prevention strategies in Toronto."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

@@

# Create a population density map with shooting crime count
ggplot(crime_data_sf) +
  geom_sf(aes(fill = POPULATION_2023), color = "black", alpha = 0.9) +
  scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +
  geom_jitter(
    data = shooting_data, aes(geometry = geometry, size = Crime_Count),
    color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002
  ) +
  scale_size_continuous(name = "Shooting Count", range = c(1, 10), breaks = c(5, 10, 15, 20)) +
  labs(
    title = "Shooting Count and Population Density",
    subtitle = "By Neighborhood",
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    legend.position = "bottom",
    legend.title = element_text(face = "bold")
  ) +
  guides(size = guide_legend(override.aes = list(alpha = 1), position = "bottom"))
```

@@

Socioeconomic factors, together with population density, play a key role in shaping crime patterns. Economically disadvantaged areas, such as Moss Park and York University Heights, tend to have higher crime rates. This underscores the importance of addressing socioeconomic disparities to reduce crime. Targeted social interventions—such as improving access to education, healthcare, and employment—could help tackle the root causes of crime.

>>>>

---
title: "Assessing Population Density and Crime Patterns in Toronto Neighborhoods (2014–2023)"
author: 
  - Yongqi Liu
thanks: "Code and data are available at:https://github.com/Cassieliu77/Crime-Patterns-and-Population-Density.git"
date: 27 September 2024
date-format: long
toc: true 
abstract: "This paper examines the relationship between population density and crime patterns across Toronto neighborhoods from 2014 to 2023. The analysis reveals that high-density areas experience higher crime rates, particularly for auto theft and assault, with socioeconomic factors further influencing crime patterns. Based on these findings, we propose targeted interventions and offer data-driven suggestions for policymakers and urban planners to enhance crime prevention strategies in Toronto."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

In the sections that follow, this paper outlines the data used and measurement methods, presents the results derived from the data and discusses the implications of these results. These findings underscore the critical need for nuanced urban planning and tailored law enforcement strategies to mitigate crime in densely populated areas of Toronto.

@@

# Create a population density map with shooting crime count
ggplot(crime_data_sf) +
  geom_sf(aes(fill = POPULATION_2023), color = "black", alpha = 0.9) +
  scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +
  geom_jitter(
    data = shooting_data, aes(geometry = geometry, size = Crime_Count),
    color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002
  ) +
  scale_size_continuous(name = "Shooting Count", range = c(1, 10), breaks = c(5, 10, 15, 20)) +
  labs(
    title = "Shooting Count and Population Density",
    subtitle = "By Neighborhoods",
    x = "Longitude", y = "Latitude"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),
    legend.position = "bottom",
    legend.title = element_text(face = "bold")
  ) +
  guides(size = guide_legend(override.aes = list(alpha = 1), position = "bottom"))
```

@@

Socioeconomic factors, together with population density, play a key role in shaping crime patterns. Economically disadvantaged areas, such as Moss Park and York University Heights, tend to have higher crime rates. This underscores the importance of addressing socioeconomic disparities to reduce crime. Targeted social interventions such as improving access to education, healthcare, and employment could help tackle the root causes of crime.
